### 2014-08-11 Docs/flowchart/flowchart.tex
more work on getting this up to date

In practice, we calculate $w_0$ by integrating
the one-dimensional divergence constraint:
\begin{equation}
\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
\end{equation}
Then we define \MarginPar{this term needs the $\beta_0$'s as shown.  Also, is this really
the form, or should it be $\overline{(\beta_0/\rho_0) \partial (\pi_0/\beta_0)/\partial r}$?}
\begin{equation}
- \frac{\beta_0}{\rho_0} \frac{\partial (\pizero/\beta_0)}{\partial r} = \frac{\partial w_0}{\partial t} + 
   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
Then we can include this for completeness in the update for $\ut.$

@@

At times, we will define a temperature equation by writing $h = h(T,p,X_k)$
and differentiating:
\begin{equation}
\label{eq:tempenthalpypsi}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left\{ \left(1 - \rho h_p\right) \left 
  [ \psi + (\Ubt \cdotb \er) \frac{\partial p_0}{\partial r} \right ]
 - \sum_k \rho \xi_k {\omegadot}_k
 + \rho \Hnuc + \rho \Hext \right \} \enskip  .
\end{equation}

@@

\begin{itemize}
\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp},                                                      
  p_0^{\inp}] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp},                                   
  T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
  evolves the species and enthalpy due to reactions through
  $\Delta t/2$ according to:
\begin{equation}
\frac{dX_k}{dt} = \omegadot_k(\rho,X_k,T) ; \qquad
\frac{dT}{dt}   = \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
\end{equation}
%This system is solved using the stiff ordinary differential equation                                     
%integration package, {\tt VODE} \citep{vode}.                                                            
Full details of the
solution procedure can be found in Paper III. We then define:
\begin{eqnarray}
(\rho\omegadot_k)^{\outp} &=& \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\dt/2}, \\
(\rho h)^{\outp} &=& (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\in\
p}.
\end{eqnarray}
where the enthalpy update includes external heat sources $(\rho\Hext)^{\inp}$.
As introduced in Paper IV, we update the temperature using $T^{\outp} =                                   
T(\rho^\outp,h^\outp,X_k^\outp)$ for planar geometry or $T^{\outp} =                                      
T(\rho^\outp,p_0^\inp,X_k^\outp)$ for spherical geometry.
Note that the density remains unchanged within {\bf React State}, i.e.,
$\rho^{\outp} = \rho^{\inp}$.

\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp] \rightarrow
  [\rho_0^\outp, \rho_0^{\outp,\nph}]$ is the process by which we
  update the base state density through $\dt$ in time.  We keep the
  time-centered edge states, $\rho_0^{\outp,\nph}$,
  since they are used later in discretization of $\etarho$ for planar problems.
\begin{description}
\item[planar:] We discretize equation (\ref{eq:Base State Density}) to
  compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} \left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ].
\end{equation}
  We compute the time-centered edge states, $\rho_0^{\outp,\nph}$,
  by discretizing an expanded form of equation (\ref{eq:Base State Density}):
\begin{equation}
\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r},
\end{equation}
  where the right hand side is used as the force term.
\item[spherical:] The base state density update now includes the area factors in the 
  divergences:
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{1}{r_j^2} \frac{\dt}{\dr} \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right].
\end{equation}
  In order to compute the time-centered edge states, an additional geometric 
  term is added to the forcing, due to the spherical discretization of 
  (\ref{eq:Base State Density}):
\begin{equation}
\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r} - \frac{2 \rho_0 w_0}{r}.
\end{equation}
\end{description}

@@

\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$
  is the process by which we update the base state enthalpy through $\dt$ in time.  
\begin{description}
\item[planar:] We discretize equation (\ref{eq:Base State Enthalpy}), neglecting reaction source terms,
  to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r} \left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} + \dt\psi_j^{\inp}.
\end{equation}
  We compute the time-centered edge states, $(\rho h)_0^{\nph}$, by discretizing
  an expanded form of equation (\ref{eq:Base State Enthalpy}):
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} + \psi.
\end{equation}
\item[spherical:]  The base state enthalpy update now includes the area factors 
  in the divergences:
\begin{eqnarray}
(\rho h)_{0,j}^{\outp} &=& (\rho h)_{0,j}^{\inp} \nonumber \\
&& - \frac{1}{r_j^2} \frac{\dt}{\dr} \left \{ \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} +\dt\psi^{\inp,\nph}.\nonumber\\
\end{eqnarray}
  In order to compute the time-centered edge states, an additional geometric 
  term is added to the forcing, due to the spherical discretization of 
  (\ref{eq:Base State Enthalpy}):
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} - \frac{2 (\rho h)_0 w_0}{r} + \psi.
\end{equation}
\end{description}

@@

In Paper III, we showed that $\psi=\etarho g$ for planar geometries, and 
derived an alternate expression for equation (\ref{eq:w0 divergence}).
We compute $w_0$ using equation (35) in Paper III using the following 
discretization:
\begin{equation}
\frac{w_{0,j+\myhalf}^\outp-w_{0,j-\myhalf}^\outp}{\Delta r} = \left(\Sbar^{\inp} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}\right)_j,
\end{equation}
with $w_{0,-\myhalf}=0$.

@@

We begin with equation (\ref{eq:w0 divergence}) written in spherical coordinates:
\begin{equation}
\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) = \beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ).
\end{equation}
We expand the spatial derivative and recall from Paper I that
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r},
\end{equation}
giving:
\begin{equation}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) = \Sbar - \frac{1}{\gammabar p_0} \underbrace{\left( \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)}_{\psi}.\label{eq:psi def}
\end{equation}
We solve this equation for $w_0$ as described in Appendix of the multilevel paper.

>>>>

In practice, we calculate $w_0$ by integrating
the one-dimensional divergence constraint.  For a plane-parallel atmosphere, the
evolution is:
\begin{equation}
\label{eq:flow:dw0dr_planar}
\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
\end{equation}

Then we define \MarginPar{this term needs the $\beta_0$'s as shown.  Also, is this really
the form, or should it be $\overline{(\beta_0/\rho_0) \partial (\pi_0/\beta_0)/\partial r}$?}
\begin{equation}
- \frac{\beta_0}{\rho_0} \frac{\partial (\pizero/\beta_0)}{\partial r} = \frac{\partial w_0}{\partial t} + 
   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
Then we can include this for completeness in the update for $\ut.$

@@

At times, we will define a temperature equation by writing $h = h(T,p,X_k)$
and differentiating:
\begin{equation}
\label{eq:flow:temp}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left\{ \left(1 - \rho h_p\right) \left 
  [ \psi + (\Ubt \cdotb \er) \frac{\partial p_0}{\partial r} \right ]
 - \sum_k \rho \xi_k {\omegadot}_k
 + \rho \Hnuc + \rho \Hext \right \} \enskip  .
\end{equation}

@@

\begin{itemize}
\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp},                                                      
  p_0^{\inp}] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp},                                   
  T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
  evolves the species and enthalpy due to reactions through
  $\Delta t/2$ according to:
\begin{equation}
\frac{dX_k}{dt} = \omegadot_k(\rho,X_k,T) ; \qquad
\frac{dT}{dt}   = \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
\end{equation}
Here the temperature equation comes from Eq.~\ref{eq:flow:temp} with $Dp_0/Dt = 0$ for
the burning part of the evolution.

Full details of the
solution procedure can be found in Paper III. We then define:
\begin{eqnarray}
(\rho\omegadot_k)^{\outp} &=& \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\dt/2}, \\
(\rho h)^{\outp} &=& (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}.
\end{eqnarray}
where the enthalpy update includes external heat sources $(\rho\Hext)^{\inp}$.
As introduced in Paper IV, we update the temperature using $T^{\outp} = 
T(\rho^\outp,h^\outp,X_k^\outp)$ for planar geometry or $T^{\outp} =
T(\rho^\outp,p_0^\inp,X_k^\outp)$ for spherical geometry, with this behavior
controlled by \runparam{use\_tfromp}.
Note that the density remains unchanged within {\bf React State}, i.e.,
$\rho^{\outp} = \rho^{\inp}$.

\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp] \rightarrow
  [\rho_0^\outp, \rho_0^{\outp,\nph}]$ is the process by which we
  update the base state density through $\dt$ in time.  We keep the
  time-centered edge states, $\rho_0^{\outp,\nph}$,
  since they are used later in discretization of $\etarho$ for planar problems.
\begin{description}
\item[planar:] We discretize equation (\ref{eq:flow:base_density}) to
  compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} \left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ].
\end{equation}
  We compute the time-centered edge states, ${\rho_0}^{\outp,\nph}_{j\pm\myhalf}$,
  by discretizing an expanded form of equation (\ref{eq:flow:base_density}):
\begin{equation}
\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r},
\end{equation}
  where the right hand side is used as the force term.
\item[spherical:] The base state density update now includes the area factors in the 
  divergences:
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{1}{r_j^2} \frac{\dt}{\dr} \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right].
\end{equation}
  In order to compute the time-centered edge states, an additional geometric 
  term is added to the forcing, due to the spherical discretization of 
  (\ref{eq:flow:base_density}):
\begin{equation}
\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r} - \frac{2 \rho_0 w_0}{r}.
\end{equation}
\end{description}

@@

\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$
  is the process by which we update the base state enthalpy through $\dt$ in time.  
\begin{description}
\item[planar:] We discretize equation (\ref{eq:flow:enthalpy_base}), neglecting reaction 
source terms, to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r} \left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} + \dt\psi_j^{\inp}.
\end{equation}
  We compute the time-centered edge states, $(\rho h)_0^{\nph}$, by discretizing
  an expanded form of equation (\ref{eq:flow:enthalpy_base}):
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} + \psi.
\end{equation}
\item[spherical:]  The base state enthalpy update now includes the area factors 
  in the divergences:
\begin{eqnarray}
(\rho h)_{0,j}^{\outp} &=& (\rho h)_{0,j}^{\inp} \nonumber \\
&& - \frac{1}{r_j^2} \frac{\dt}{\dr} \left \{ \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} +\dt\psi^{\inp,\nph}.\nonumber\\
\end{eqnarray}
  In order to compute the time-centered edge states, an additional geometric 
  term is added to the forcing, due to the spherical discretization of 
  (\ref{eq:flow:enthalpy_base}):
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} - \frac{2 (\rho h)_0 w_0}{r} + \psi.
\end{equation}
\end{description}

@@

In Paper III, we showed that $\psi=\etarho g$ for planar geometries,
and derived derived Eq.~\ref{eq:flow:dw0dr_planar} as an alternate
expression for Eq.~\ref{eq:w0 divergence}.  We discretize this as:
\begin{equation}
\frac{w_{0,j+\myhalf}^\outp-w_{0,j-\myhalf}^\outp}{\Delta r} = \left(\Sbar^{\inp} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}\right)_j,
\end{equation}
with $w_{0,-\myhalf}=0$.

@@

We begin with equation (\ref{eq:w0 divergence}) written in spherical coordinates:
\begin{equation}
\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) = \beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ).
\end{equation}
We expand the spatial derivative and recall from Paper I that
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r},
\end{equation}
giving:
\begin{equation}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) = \Sbar - \frac{1}{\gammabar p_0} \underbrace{\left( \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)}_{\psi}.\label{eq:psi def}
\end{equation}
We solve this equation for $w_0$ as described in Appendix~B of the multilevel paper.
